### 2020-05-07 paper.tex
War Stories

In order to answer RQ1 and get an insight in how software engineers apply RP, we started with an early series of informal interviews. The only requirement to qualify for the interview was to apply RP in some way, hence no restriction in regards of tech stack was given. In the end we interviewed five engineers: Four work or worked more recently with Angular\footnote{\url{todo}} and ngrx\footnote{\url{todo}} to develop frontend applications. Two of them stated to have more explicit experience with RxJS further. The fifth subject works currently with a combination of akka\footnote{\url{todo}} and akka-streams\footnote{\url{todo}} in Scala to create backend applications. The agenda available in Appendix~\ref{sec:interviewquestions} was used to maintain focus during the interview sessions.

@@

None of the subjects remembered debugging RP code to be enjoyable in particular, nonetheless the technology used to do so. All of the subjects use traditional debugging tools. Redux DevTools \footnote{\url{todo}} is in use by all four subjects working with ngrx. There was mention of any more specific debugging tool in regards of akka and akka-streams.

One particular subject with more in-depth knowledge on RxJS described situations where they would have wished for tools which could ``understand'' the actual \mintinline{TypeScript}{Observable} graph better: (i) When an error is thrown within the function passed to an operator, the generated stack trace reflects where the operator called the function rather than the actual location where the function is implemented. (ii) In order to understand how data traverses through the graph, subjects augmented code manually with the \mintinline{TypeScript}{tap}\footnote{\url{TODO}} operator to log values. A less invasive way would be preferable as stated.

@@

>>>>

In order to prepare to give an answer on the first research question, we started with a series of informal interviews. We talked to professionals about their experience in working with RP in their daily jobs. We did not limit the selection of our interview partners by the tech stack defined through our research area since we were interested in as many different perspective on the topic as possible in this early stage. Eventually, we interviewed five engineers: Four work or worked more recently with Angular\footnote{\url{todo}} and ngrx\footnote{\url{todo}} to develop frontend applications. Two of them stated to have more explicit experience with RxJS further. The fifth subject works currently with a combination of akka\footnote{\url{todo}} and akka-streams\footnote{\url{todo}} in Scala to create backend applications. The agenda available in Appendix~\ref{sec:interviewquestions} was used to maintain focus during the interview sessions.

@@

None of the subjects remembered debugging RP code to be a positive experience, independent of the technology used. All of the subjects use traditional debugging tools. Redux DevTools \footnote{\url{todo}} is in use by all four subjects working with ngrx. There was mention of any more specific debugging tool in regards of akka and akka-streams.

One particular subject with more in-depth knowledge on RxJS described situations where they would have wished for tools which could ``understand'' the actual \mintinline{TypeScript}{Observable} graph better: (i) When an error is thrown within the function passed to an operator, the generated stack trace reflects where the operator called the function rather than the actual location where the function is implemented. (ii) In order to understand how data traverses through the graph, subjects augmented code manually with the \mintinline{TypeScript}{tap}\footnote{\url{todo}} operator to log values. A less invasive way would be preferable as stated.

@@

Based on the results of the interviews, we compiled a request for more specific factual reports. We asked professional software engineers about their most recent debugging experience in regards of RxJS, either good or bad. We were able to gather responses from one RxJS core team member, two Google Developer Experts and two engineers building web and mobile applications for a fintech company eventually.

We could verify gathered interview data with the results from the new reports: Subjects tend to add log statements to their code in order to understand its behavior, especially when things get more complex. Three subjects even claimed using specialized logging functionalities provided through rxjs-spy\footnote{url{todo}}. This library allows to ``tag'' specific points in the \mintinline{TypeScript}{Observable} graph with a probe. In addition to report values traversing the graph, each probe notifies about events when an observer subscribes or unsubscribes to them or when the source \mintinline{TypeScript}{Observable} completes \cite{rxjs-lifecycle} \cite{rxjs-spy-events}. As the subjects stated, the knowledge about these additional events helps in situations where they need to find problems within a graph composing multiple asynchronous data sources on multiple levels. Without it, they were blind to what actually happens within the system.

In addition to more sophisticated logging, three subjects reported that they built a habit of decomposing complex graphs into smaller, simpler bits and running those in a dedicated environment like Rx Visualizer\footnote{\url{https://rxviz.com/}} or StackBlitz\footnote{\url{https://stackblitz.com/}}.
